Thread Pool in GeckoDB/BOLSTER - Integrating a Thread Pool into a high performance Database System \\

### 2018-06-19 Presentations/Paper/paper_thread_pool.tex
Add first version of abstract and env to paper

\author{

	\IEEEauthorblockN{Robert Jendersie, Johannes Wuensche, Johann Wagner, Marten Wallewein-Eising}
	\IEEEauthorblockA{\textit{Otto-von-Guericke University} \\
		Magdeburg, Germany \\
		\emph{firstname}.\emph{lastname}@st.ovgu.de} \and

}

@@

\begin{abstract}

\end{abstract}

@@

\subsection{Managing Thread States}
As mentioned before, accessing the same memory location from different thread may lead to race conditions, which means the program behaviour depends on the accessing order of the threads. For example, while one thread updates data at a specific memory location another threads writes into the data, which results to inconsistency. In Section 2, we presented two approaches to handle multithreaded access to the same memory location, mutex threads and atomic operations and we state where those approaches are used in our implementation. In this subsection, we focus on atomic operations to manage thread states. We consider the following steps for a thread to update its state. At first, the current state has to be checked and, depending on the current state, the state has to be updated. If these steps are implemented using an if statement and an update function, the complete update procedure is not atomic. Consequently, another thread can update the state between the check of the current state and the update. To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} of the stdatomic lib. This function combines the check and update in one atomic procedure. We implemented every thread state access with this function to ensure atomic thread state updates. After presenting the enqueueing, processing, and waiting for tasks, we show the second component of our thread pool, the performance monitoring.

@@

\section{Experimental Environment}

@@

\section{Conclusion and Future Work}
\cite{pyarali2001evaluating}
\bibliographystyle{splncs}
\bibliography{paper_thread_pool}

>>>>

\author{
	\IEEEauthorblockN{Robert Jendersie, Johannes Wuensche, Johann Wagner, Marten Wallewein-Eising}
	\IEEEauthorblockA{\textit{Otto-von-Guericke University} \\
		Magdeburg, Germany \\
		\emph{firstname}.\emph{lastname}@st.ovgu.de} \and
}

@@

\begin{abstract}
Processing tasks in parallel is used in nearly all applications to keep up with the requirements on modern software systems. However, the implementation of parallel processing can differ strongly depending on the desired use cases and reach from spawning dozens of threads to reuse threads in thread pools. In this paper, we show an implementation of a thread pool to process independent tasks in parallel including the waiting for a set of processed tasks. We also compare the thread pool implementation against a baseline, which spawns new threads for parallel data processing. The evaluation of the thread pool shows, that designing and implementing a thread pool is a balancing act between generalization of tasks and the processing performance. Additionally, we show that the task and thread pool configuration depending on the use case has a high impact on the thread pool performance.
\end{abstract}

@@

\subsection{Managing Thread States}
As mentioned before, accessing the same memory location from different thread may lead to race conditions, which means the program behaviour depends on the accessing order of the threads. For example, while one thread updates data at a specific memory location another threads writes into the data, which results to inconsistency. In Section 2, we presented two approaches to handle multithreaded access to the same memory location, mutex threads and atomic operations and we state where those approaches are used in our implementation. In this subsection, we focus on atomic operations to manage thread states. We consider the following steps for a thread to update its state. At first, the current state has to be checked and, depending on the current state, the state has to be updated. If these steps are implemented using an if statement and an update function, the complete update procedure is not atomic. Consequently, another thread can update the state between the check of the current state and the update. To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} \cite{atomicstrong} of the stdatomic lib. This function combines the check and update in one atomic procedure. We implemented every thread state access with this function to ensure atomic thread state updates. After presenting the enqueueing, processing, and waiting for tasks, we show the second component of our thread pool, the performance monitoring.

@@

\section{Experimental Environment}	
In this section, we present our experimental environment focussing on the design of our benchmarks and the system configuration we used to evaluate our thread pool implementation.

\subsection{Evaluation Setup}
In our evaluation setup, we focus primarily on the comparison of using our thread pool against the baseline implementation of creating new threads at runtime of BOLSTER. We simulate calls of BOLSTER primitives with different thread pool and task configurations and compare the results against the baseline.

Additionally, we evaluate the metrics for threads, tasks and the thread pool mentioned in Section 3.F regarding the thread pool and task configuration used in the BOLSTER primitives. To evaluate optimal configurations for the thread pool, we consider the relation between idle and busy time of the thread pool and the average waiting time of tasks.

Furthermore, we compare the scope and usage of our thread pool against an existing implementation \emph{Threading Building Blocks (Intel TBB)} \textbf{TODO: cite!} written in C++. Since Intel TBB uses completely different approaches to parallel the execution of tasks, a direct performance comparison may not lead to reliable results. Therefore, we evaluate the usage and the scope both implementations give.

\subsection{Experimental Configuration}
\textbf{TODO: Change?}\\The operating system is Ubuntu 17.04 with linux kernel version 4.13 running in a virtual machine on an Intel Core i7-6700HQ, 4x 2.6 GHz (Skylake) and 8GB DDR3 RAM running on 1600Mhz. We reduce the number of cores of the virtual machine to 4 instead of 8 virtual cores of the CPU. The CPU has 256KB L1 cache, 1MB L2 cache, and 6MB L3 cache. All simulations are repeated three times and a median value of all repetitions is used for evaluation.

@@

\section{Conclusion and Future Work}

\bibliographystyle{splncs}
\bibliography{paper_thread_pool}
